---
title: "HW 4"
author: "Caitlin Bolz"
date: "3/11/2021"
output: html_document
---

```{r}
library(shiny)
library(maps)
library(mapproj)
```

Go to github.com. Make an empty repository called CensusApp. Clone the repo in Rstudio. Then, go through Lesson 5 to construct the app in that repository. Commit and push your repository. Then, turn in the command


```{r}
source("CensusApp/helper.R")
counties <- readRDS("CensusApp/data/counties.rds")
percent_map(counties$white, "darkgreen", "% White")
```


```{r}
shinyApp(ui, server)
```






# Runs your code
```{r}
runGitHub("CensusApp", "cabolz")
```






# User interface ----
ui <- fluidPage(
  titlePanel("censusVis"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Create demographic maps with 
        information from the 2010 US Census."),
      
      selectInput("var", 
                  label = "Choose a variable to display",
                  choices = c("Percent White", "Percent Black",
                              "Percent Hispanic", "Percent Asian"),
                  selected = "Percent White"),
      
      sliderInput("range", 
                  label = "Range of interest:",
                  min = 0, max = 100, value = c(0, 100))
    ),
    
    mainPanel(plotOutput("map"))
  )
)

# Server logic ----
```{r}
server <- function(input, output) {
  output$map <- renderPlot({
    data <- switch(input$var, 
                   "Percent White" = counties$white,
                   "Percent Black" = counties$black,
                   "Percent Hispanic" = counties$hispanic,
                   "Percent Asian" = counties$asian)
    
    color <- switch(input$var, 
                    "Percent White" = "darkgreen",
                    "Percent Black" = "black",
                    "Percent Hispanic" = "darkorange",
                    "Percent Asian" = "darkviolet")
    
    legend <- switch(input$var, 
                     "Percent White" = "% White",
                     "Percent Black" = "% Black",
                     "Percent Hispanic" = "% Hispanic",
                     "Percent Asian" = "% Asian")
    
    percent_map(data, color, legend, input$range[1], input$range[2])
  })
}
```
# Run app ----
shinyApp(ui, server)