---
title: "STAT 433 HW 1"
author: "Caitlin Bolz"
date: "2/3/2021"
output: github_document
header-includes:
    - \usepackage {hyperref}
    - \hypersetup {colorlinks = true, linkcolor = red, urlcolor = red}
---

My GitHub repository for this class can be found at this [link](https://github.com/cabolz/STAT433-HW) 
```{r}
library(readr)
library(data.table)
```

I chose a smaller dataset to work with so I could easily read and interpret the file. Once I apply my findings, this code can be used for other bridge data sets by simply changing the link in the read.csv command.

The specific data set I used can be found at this [link](https://www.fhwa.dot.gov/bridge/nbi/2019/delimited/AK19.txt)

I initially started reading in the data with the read.csv, read_csv, and fread commands. All three of these commands produced 1595 obs. of 123 variables. There are 1,595 bridges in Alaska and 123 variables being measured so this is correct. I did have to use header = T for the read.csv

```{r}
bridgesA = read.csv("https://www.fhwa.dot.gov/bridge/nbi/2019/delimited/AK19.txt", header = T)

bridgesB = read.csv("https://www.fhwa.dot.gov/bridge/nbi/2019/delimited/AK19.txt")

bridgesC = fread("https://www.fhwa.dot.gov/bridge/nbi/2019/delimited/AK19.txt")
```


Then looked at the head of each dataset and their structures. I found that the heads were all the same. I examined the first 60 or so variables and the strucutre was the same for each column across the three sets.
```{r}
head(bridgesA, n = c(3L, 6L))
head(bridgesB, n = c(3L, 6L))
head(bridgesC, n = c(2L))

str(bridgesA)
str(bridgesB)
str(bridgesC)


```

