---
title: "Week 5 In Class Review Exercises"
author: "Caitlin Bolz"
date: "2/25/2021"
output: html_document
---


```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(nycflights13)
```

# Question 1
Make a table that describes each plane. It should have a column for tailnum, another column for average arrival delay, and another for the year the plane was manufactured.

```{r}
pr1 = flights %>% 
  group_by(tailnum) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = T)) %>% 
  left_join(planes) %>% 
  select(tailnum:year)

head(pr1, n = 10)
```


    
# Question 2
Make a table where each row is a day of the year. The first column is the date. The 2:4 columns give the number of (scheduled) departures from EWR, LGA, and JFK.

```{r}
pr2 = flights %>% 
  mutate(date = as.Date(time_hour)) %>% 
  group_by(origin, date) %>% 
  summarise(deps = n(), .groups = 'keep') %>% 
  pivot_wider(names_from = origin, values_from = deps)

head(pr2, n = 10)
```



# Question 3
Make a table where each row is a day of the year. Each destination airport is a column. The elements (day x destination) give the number of flights to that destination. What should NA’s be?
```{r}
pr3 = flights %>% 
  mutate(date = as.Date(time_hour)) %>% 
  group_by(date, dest) %>% 
  summarise(num_arr = sum(!is.na(arr_time)), .groups = 'keep') %>% 
  pivot_wider(names_from = dest, values_from = num_arr, values_fill = list(num_arr = 0))

head(pr3, n = 10)
```