---
title: "STAT 433 HW2"
author: "Caitlin Bolz"
date: "2/10/2021"
output: github_document
header-includes:
    - \usepackage {hyperref}
    - \hypersetup {colorlinks = true, linkcolor = red, urlcolor = red}
---


```{r}
library(dplyr)
library(nycflights13)
library(ggplot2)
```


# Question 1
How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

8,255 rows with missing dep_time. They also have missing dep_delay, arr_time, arr_delay, and air_time. These rows represent flights that never took off, so canceled flights.
```{r}
flights %>% filter(is.na(dep_time))
```


# Question 2
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights$dep_time = (floor(flights$dep_time/100)*60 + flights$dep_time %% 100)

flights$sched_dep_time = (floor(flights$sched_dep_time/100)*60 + 
                            flights$sched_dep_time %% 100)
```



# Question 3
Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with `ggplot(aes(x= ,y=)) + geom_point()’


```{r}
flights %>%
  group_by(day) %>%
  summarise(cancelled = mean(is.na(dep_delay)),
            mean_dep = mean(dep_delay, na.rm = T),
            mean_arr = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = cancelled)) +
  geom_point(aes(x = mean_dep), colour = "red") +
  geom_point(aes(x = mean_arr), colour = "blue") +
  labs(x = "Avg Delay per Day", y = "Proportion of Cancelled Flights per Day", title = "Average Delay vs Cancelled Flights")
```